07_species_layers

### 2025-10-21 05_formulating_feeds.qmd
New novel feeds with digestibility

The purpose of this markdown is to:

@@

here("src") %>% list.files(pattern = "\\.R$", full.names = TRUE) %>% purrr::walk(source)

ingred_inputs_file <- file.path(input_feed_profile_path, "all_ingredients.csv")
feed_input_file <- file.path(input_feed_profile_path, "all_feeds.xlsx")
```

@@

There is no check or correction here to ensure that protein+lipid+ash+carb = 1 in the incoming data. This was checked manually prior to import.

``` {r import-ingredients}
#| code-summary: Ingredient details

ingreds <- ingred_inputs_file %>% 
  read.csv() %>% 
  mutate(ingredient = as.factor(ingredient))
ingred_nms <- levels(ingreds$ingredient)
```

@@

feed_inputs <- feed_input_file %>% 
  read_excel(sheet = "all_feeds") %>% 
  pivot_longer(names_to = "feed", values_to = "proportion", cols = !contains(c("ingredient", "category"))) %>% 
  mutate(feed = as.factor(feed),
         ingredient = as.factor(ingredient)) %>% 
  merge(ingreds, by = "ingredient")

@@

feed_inputs <- feed_inputs %>% 
  filter(!is.na(proportion) & proportion != 0) %>% 
  group_by(feed) %>% 
  mutate(
    sum = sum(proportion),
    proportion = proportion/sum
  ) %>% 
    ungroup()
```

@@

feed_params <- purrr::map(feed_types, function(ft) {
  df <- feed_inputs %>% 
    filter(feed == ft) 
  list(
    Proteins = df %>% 
      select(ingredient, proportion, contains("protein"), -contains("feed")) %>%
      rename(macro = protein, digest = protein_digestibility),
    Carbohydrates = df %>% 
      select(ingredient, proportion, contains("carb"), -contains("feed")) %>%
      rename(macro = carb, digest = carb_digestibility),
    Lipids = df %>% 
      select(ingredient, proportion, contains("lipid"), -contains("feed")) %>%
      rename(macro = lipid, digest = lipid_digestibility)
  )
}) %>% 
  setNames(feed_types)
```

>>>>

The purpose of this markdown is to formulate all the feeds that will be used in the analysis. These feeds are:
- marine_dominant: a "past" salmonid diet where most protein is sourced from marine-based ingredients.
- plant_dominant: a "current" salmonid diet where most protein is sourced from land-based ingredients.
- novel_inclusive_2: a potential future salmonid diet which includes some novel ingredients. This formulation uses maximal digestibility values.
- novel_inclusive_1: a potential future salmonid diet which includes some novel ingredients. This formulation uses the same formulation as novel_inclusive_2, but with minimal digestibility values.

@@

here("src") %>% list.files(pattern = "\\.R$", full.names = TRUE) %>% purrr::walk(source)
```

@@

There is no check or correction here to ensure that protein + lipid + ash + carb = 1 in the incoming data. This was checked manually prior to import.

``` {r import-ingredients}
#| code-summary: Get ingredient details (PLC composition, digestibility)

ingreds <- file.path(input_feed_profile_path, "all_ingredients.xlsx") %>% 
  read_excel(sheet = "all_ingredients") %>% 
  mutate(ingredient = as.factor(ingredient))

ingred_nms <- levels(ingreds$ingredient)
```

@@

feed_inputs <- file.path(input_feed_profile_path, "all_feeds.xlsx") %>% 
  read_excel(sheet = "all_feeds") %>% 
  rename(novel_inclusive_1 = novel_inclusive)
feed_inputs$novel_inclusive_2 <- feed_inputs$novel_inclusive_1 # same composition

feed_inputs <- feed_inputs %>% 
  pivot_longer(names_to = "feed", values_to = "proportion", cols = !contains(c("ingredient", "category"))) %>% 
  mutate(
    feed = as.factor(feed),
    ingredient = as.factor(ingredient),
    proportion = case_when(is.na(proportion) ~ 0, T ~ proportion)
  ) %>% 
  left_join(ingreds, by = "ingredient") %>% 
  select(-prop_P, -P_digestibility)

# Novel feeds use the minimum and maximum digestibility values
feed_inputs <- feed_inputs %>% 
  mutate(
    protein_digestibility = case_when(
      feed == "novel_inclusive_1" & !is.na(max_protein_digestibility) ~ max_protein_digestibility,
      feed == "novel_inclusive_2" & !is.na(min_protein_digestibility) ~ min_protein_digestibility,
      T ~ mean_protein_digestibility,
    ),
    carb_digestibility = case_when(
      feed == "novel_inclusive_1" & !is.na(max_carb_digestibility) ~ max_carb_digestibility,
      feed == "novel_inclusive_2" & !is.na(min_carb_digestibility) ~ min_carb_digestibility,
      T ~ mean_carb_digestibility,
    ),
    lipid_digestibility = case_when(
      feed == "novel_inclusive_1" & !is.na(max_lipid_digestibility) ~ max_lipid_digestibility,
      feed == "novel_inclusive_2" & !is.na(min_lipid_digestibility) ~ min_lipid_digestibility,
      T ~ mean_lipid_digestibility,
    )
  ) %>% 
  select(-contains(c("min", "max", "mean")))

@@

feed_inputs <- feed_inputs %>% 
  filter(!is.na(proportion) & proportion != 0) %>% 
  group_by(feed) %>% 
  mutate(
    sum = sum(proportion),
    proportion = proportion/sum
  ) %>% 
  ungroup() %>% 
  select(-sum, -ash)
```

@@

feed_params <- purrr::map(feed_types, function(ft) {
  df <- feed_inputs %>% 
    filter(feed == ft) 
  list(
    Proteins = df %>% 
      select(ingredient, proportion, contains("protein")) %>% 
      rename(macro = protein, digest = protein_digestibility),
    Carbohydrates = df %>% 
      select(ingredient, proportion, contains("carb")) %>% 
      rename(macro = carb, digest = carb_digestibility),
    Lipids = df %>% 
      select(ingredient, proportion, contains("lipid")) %>% 
      rename(macro = lipid, digest = lipid_digestibility)
  )
}) %>% 
  setNames(feed_types)
```

### 2025-10-21 06_run_atlantic_salmon_model.qmd
New novel feeds with digestibility

# Check that the feed will actually work
targets::tar_make(test_reference_feed)
targets::tar_load(test_reference_feed)
test_reference_feed # if weight = 0 then the feeds aren't processing properly

@@

targets::tar_validate()
# targets::tar_visnetwork(targets_only = T)
# targets::tar_outdated()
targets::tar_make(
  seconds_meta_append = 120,
  callr_function = NULL
  )
tar_prune()
```

@@

Note that this takes a long time - each branch of farmrun takes ~5.3 minutes, and a maximum of ~12 workers can run at a time. The total therefore takes ~17 hours.

@@

# Pre-run to reduce load
targets::tar_make(
  names = c("farm_IDs_chunked", "farm_ts_data_chunked", "farm_IDs", "test_reference_feed", "farm_temp_data_chunked", "harvest_size_chunked", "farm_static_data_chunked"), 
  seconds_meta_append = 90#, callr_function = NULL
)

# Run results
targets::tar_make(
  names = c("farm_run_chunked", "stat_names", "farm_results_chunked", "biomass_produced_chunked", "cohort_results_chunked" ), 
  seconds_meta_append = 90,
  shortcut = T
)

>>>>

Sys.setenv(TAR_PROJECT = "project_farmruns")
tar_prune() # will remove any targets objects no longer in pipeline

# Check that the feed will actually work
targets::tar_make(test_reference_feed)
targets::tar_load(test_reference_feed)
test_reference_feed # if weight ~ 0 then the feeds aren't processing properly

@@

targets::tar_validate()
# targets::tar_visnetwork(targets_only = T)
# targets::tar_outdated()

targets::tar_make(
  seconds_meta_append = 120,
  callr_function = NULL
  )
tar_prune()
```

@@

@@

# Pre-run to reduce load
targets::tar_make(
  names = c("farm_static_data_chunked", "farm_temp_data_chunked", "farm_ts_data_chunked"), 
  seconds_meta_append = 90
)

# Run results
targets::tar_make(
  names = c("farm_run_chunked", "farm_full_results"), 
  seconds_meta_append = 90,
  shortcut = T # only if you've done the pre-run, above
)

### 2025-10-21 07_species_layers.qmd
Proper index of files
New novel feeds with digestibility

library(tidyr)
library(dplyr)
library(magrittr)
library(stringr)
library(here)
library(terra)
library(sf)
library(qs)
library(ggplot2)
library(countrycode)
library(terra)
library(tidyterra)
library(cowplot)
library(furrr)
library(rnaturalearth)
library(aquamapsdata) # https://raquamaps.github.io/aquamapsdata/articles/intro.html
library(worrms)
library(grid)

source("00_dirs.R")

overwrite <- F
```

>>>>

library(tidyr)
library(dplyr)
library(magrittr)
library(stringr)
library(here)
library(terra)
library(sf)
library(qs)
library(ggplot2)
library(countrycode)
library(terra)
library(tidyterra)
library(cowplot)
library(furrr)
library(rnaturalearth)
# library(aquamapsdata) # https://raquamaps.github.io/aquamapsdata/articles/intro.html
library(worrms)
library(grid)

here("src") %>% list.files(pattern = "\\.R$", full.names = TRUE) %>% walk(source)
```
